
  \begin{itemize}[leftmargin=*]
\item \textbf{Legendre symbol:} $\left(\frac{a}{b}\right) = a^{(b-1)/2} \pmod{b}$, $b$ odd prime.
    \item \textbf{Heron's formula:} A triangle with side lengths
      $a,b,c$ has area $\sqrt{s(s-a)(s-b)(s-c)}$ where $s =
      \frac{a+b+c}{2}$.
    \item \textbf{Pick's theorem:} A polygon on an integer grid
      strictly containing $i$ lattice points and having $b$ lattice
      points on the boundary has area $i + \frac{b}{2} - 1$. (Nothing
      similar in higher dimensions)
    \item \textbf{Euler's totient:} The number of integers less than
      $n$ that are coprime to $n$ are $n\prod_{p|n}\left(1 - \frac{1}{p}\right)$
      where each $p$ is a distinct prime factor of $n$.
    \item \textbf{König's theorem:} In any bipartite graph $G=(L\cup R,E)$, the number
      of edges in a maximum matching is equal to the number of
      vertices in a minimum vertex cover. Let $U$ be the set of
      unmatched vertices in $L$, and $Z$ be the set of vertices that
      are either in $U$ or are connected to $U$ by an alternating
      path. Then $K=(L\setminus Z)\cup(R\cap Z)$ is the minimum
      vertex cover.
    \item A minumum Steiner tree for $n$ vertices requires at most $n-2$ additional Steiner vertices.
    \item The number of vertices of a graph is equal to its minimum
      vertex cover number plus the size of a maximum independent set.
    \item \textbf{Lagrange polynomial} through points $(x_0,y_0),\ldots,(x_k,y_k)$ is $L(x) = \sum_{j=0}^k y_j \prod_{\shortstack{$\scriptscriptstyle 0\leq m \leq k$ \\ $\scriptscriptstyle m\neq j$}} \frac{x-x_m}{x_j - x_m}$
    \item \textbf{Hook length formula:} If $\lambda$ is a Young diagram and $h_{\lambda}(i,j)$ is the hook-length of cell $(i,j)$, then then the number of Young tableux $d_{\lambda} = n!/\prod h_{\lambda}(i,j)$.
    \item \textbf{Möbius inversion formula:} If $f(n) = \sum_{d|n} g(d)$, then $g(n) = \sum_{d|n} \mu(d) f(n/d)$. If $f(n) = \sum_{m=1}^n g(\lfloor n/m\rfloor)$, then $g(n) = \sum_{m=1}^n \mu(m)f(\lfloor\frac{n}{m}\rfloor)$.
    \item \#primitive pythagorean triples with hypotenuse $<n$ approx $n/(2\pi)$.
    \item \textbf{Frobenius Number:} largest number which can't be
      expressed as a linear combination of numbers $a_1,\ldots,a_n$
      with non-negative coefficients. $g(a_1,a_2) = a_1a_2-a_1-a_2$,
      $N(a_1,a_2)=(a_1-1)(a_2-1)/2$. $g(d\cdot a_1,d\cdot a_2,a_3) =
      d\cdot g(a_1,a_2,a_3) + a_3(d-1)$. An integer $x>\left(\max_i
      a_i\right)^2$ can be expressed in such a way iff.\ $x\ |\
      \mathrm{gcd}(a_1,\ldots,a_n)$.
  \end{itemize}

  \subsection{Physics}
    \begin{itemize}
      \item \textbf{Snell's law:} $\frac{\sin\theta_1}{v_1} = \frac{\sin\theta_2}{v_2}$
    \end{itemize}

  \subsection{Markov Chains}
    A Markov Chain can be represented as a weighted directed graph of
    states, where the weight of an edge represents the probability of
    transitioning over that edge in one timestep. Let $P^{(m)} = (p^{(m)}_{ij})$
    be the probability matrix of transitioning from state $i$ to state $j$
    in $m$ timesteps, and note that $P^{(1)}$ is the adjacency matrix of
    the graph. \textbf{Chapman-Kolmogorov:} $p^{(m+n)}_{ij} = \sum_{k}
    p^{(m)}_{ik} p^{(n)}_{kj}$. It follows that $P^{(m+n)} =
    P^{(m)}P^{(n)}$ and $P^{(m)} = P^m$. If $p^{(0)}$ is the initial
    probability distribution (a vector), then $p^{(0)}P^{(m)}$ is the
    probability distribution after $m$ timesteps.

    The return times of a state $i$ is $R_i = \{m\ |\ p^{(m)}_{ii} > 0 \}$,
    and $i$ is \textit{aperiodic} if $\gcd(R_i) = 1$. A MC is aperiodic if
    any of its vertices is aperiodic. A MC is \textit{irreducible} if the
    corresponding graph is strongly connected.

    A distribution $\pi$ is stationary if $\pi P = \pi$. If MC is
    irreducible then $\pi_i = 1/\mathbb{E}[T_i]$, where $T_i$ is the
    expected time between two visits at $i$. $\pi_j/\pi_i$ is the expected
    number of visits at $j$ in between two consecutive visits at $i$. A MC
    is \textit{ergodic} if $\lim_{m\to\infty} p^{(0)} P^{m} = \pi$. A MC is
    ergodic iff.\ it is irreducible and aperiodic.

    A MC for a random walk in an undirected weighted graph (unweighted
    graph can be made weighted by adding $1$-weights) has $p_{uv} =
    w_{uv}/\sum_{x} w_{ux}$. If the graph is connected, then $\pi_u =
    \sum_{x} w_{ux} / \sum_{v}\sum_{x} w_{vx}$. Such a random walk is
    aperiodic iff.\ the graph is not bipartite.

    An \textit{absorbing} MC is of the form $P = \left(\begin{matrix} Q & R
    \\ 0 & I_r \end{matrix}\right)$. Let $N = \sum_{m=0}^\infty Q^m = (I_t
    - Q)^{-1}$. Then, if starting in state $i$, the expected number of
    steps till absorption is the $i$-th entry in $N1$. If starting in state
    $i$, the probability of being absorbed in state $j$ is the $(i,j)$-th
    entry of $NR$.

    Many problems on MC can be formulated in terms of a system of
    recurrence relations, and then solved using Gaussian elimination.

  \subsection{Burnside's Lemma}
    Let $G$ be a finite group that acts on a set $X$. For each $g$ in $G$
    let $X^g$ denote the set of elements in $X$ that are fixed by $g$. Then
    the number of orbits \[ |X/G| = \frac{1}{|G|} \sum_{g\in G} |X^g| \]

    \[
        Z(S_n) = \frac{1}{n} \sum_{l=1}^n a_l Z(S_{n-l})
    \]

  \subsection{Bézout's identity}
    If $(x,y)$ is any solution to $ax+by=d$ (e.g.\ found by the Extended
    Euclidean Algorithm), then all solutions are given by \[
    \left(x+k\frac{b}{\gcd(a,b)}, y-k\frac{a}{\gcd(a,b)}\right) \]

  \subsection{Misc}
    \subsubsection{Determinants and PM}
      \begin{align*}
        det(A) &= \sum_{\sigma \in S_n}\text{sgn}(\sigma)\prod_{i = 1}^n a_{i,\sigma(i)}\\
        perm(A) &= \sum_{\sigma \in S_n} \prod_{i = 1}^n a_{i,\sigma(i)}\\
        pf(A) &= \frac{1}{2^nn!}\sum_{\sigma \in S_{2n}} \text{sgn}(\sigma)\prod_{i = 1}^n a_{\sigma(2i-1),\sigma(2i)}\\ &= \sum_{M \in \text{PM}(n)} \text{sgn}(M) \prod_{(i,j) \in M} a_{i,j}
      \end{align*}

    \subsubsection{BEST Theorem}
      Count directed Eulerian cycles. Number of OST given by
      Kirchoff's Theorem (remove r/c with root) $\#\textsc{OST}(G,r)
      \cdot \prod_{v}(d_v-1)!$

    \subsubsection{Primitive Roots}
      Only exists when $n$ is $2, 4, p^k, 2p^k$, where $p$ odd prime. Assume
      $n$ prime. Number of primitive roots $\phi(\phi(n))$
      Let $g$ be primitive root. All primitive roots are of the form $g^k$
      where $k,\phi(p)$ are coprime.\\ $k$-roots:
      $g^{i \cdot \phi(n) / k}$ for $0 \leq i < k$

    \subsubsection{Sum of primes} For any multiplicative $f$:
      \[
          S(n,p) = S(n, p-1) - f(p) \cdot (S(n/p,p-1) - S(p-1,p-1))
      \]

    \subsubsection{Floor}
      \begin{align*}
          &\left\lfloor \left\lfloor x/y \right\rfloor / z \right\rfloor = \left\lfloor x / (yz) \right\rfloor \\
          &x \% y = x - y \left\lfloor x / y \right\rfloor
      \end{align*}
<<<<<<< Updated upstream
    \subsubsection{Large Primes}
      100894373, 103893941, 999999937, 1000000007, 16208191877


\iffalse
  \clearpage
  \section*{Practice Contest Checklist}
    \begin{itemize}
      \item How many operations per second? Compare to local machine.
      \item What is the stack size?
      \item How to use printf/scanf with long long/long double?
      \item Are \texttt{\_{}\_{}int128} and \texttt{\_{}\_{}float128} available?
      \item Does MLE give RTE or MLE as a verdict? What about stack overflow?
      \item What is \texttt{RAND\_{}MAX}?
      \item How does the judge handle extra spaces (or missing newlines) in the output?
      \item Look at documentation for programming languages.
      \item Try different programming languages: C++, Java and Python.
      \item Try the submit script.
      \item Try local programs: i?python[23], factor.
      \item Try submitting with \texttt{assert(false)} and \texttt{assert(true)}.
      \item Return-value from \texttt{main}.
      \item Look for directory with sample test cases.
      \item Make sure printing works.

      \item Remove this page from the notebook.
    \end{itemize}
\fi
=======

\subsection{Maximum Density Subgraph}
    Given (weighted) undirected graph $G$. Binary search density. If $g$ is
    current density, construct flow network: $(S, u, m)$, $(u, T,
    m+2g-d_u)$, $(u,v,1)$, where $m$ is a large constant (larger than sum
    of edge weights). Run floating-point max-flow. If minimum cut has empty
    $S$-component, then maximum density is smaller than $g$, otherwise it's
    larger. Distance between valid densities is at least $1/(n(n-1))$. Edge
    case when density is $0$. This also works for weighted graphs by
    replacing $d_u$ by the weighted degree, and doing more iterations (if
    weights are not integers).
  \subsection{Maximum-Weight Closure}
    Given a vertex-weighted directed graph $G$. Turn the graph into a flow
    network, adding weight $\infty$ to each edge. Add vertices $S,T$. For
    each vertex $v$ of weight $w$, add edge $(S,v,w)$ if $w\geq 0$, or edge
    $(v,T,-w)$ if $w<0$. Sum of positive weights minus minimum $S-T$ cut is
    the answer. Vertices reachable from $S$ are in the closure. The
    maximum-weight closure is the same as the complement of the
    minimum-weight closure on the graph with edges reversed.
  \subsection{Maximum Weighted Ind. Set in a Bipartite Graph}
    This is the same as the minimum weighted vertex cover. Solve this by
    constructing a flow network with edges $(S,u,w(u))$ for $u\in L$,
    $(v,T,w(v))$ for $v\in R$ and $(u,v,\infty)$ for $(u,v)\in E$. The
    minimum $S,T$-cut is the answer. Vertices adjacent to a cut edge are
    in the vertex cover.
  \subsection{Synchronizing word problem}
    A DFA has a synchronizing word (an input sequence that moves all states
    to the same state) iff.\ each pair of states has a synchronizing word.
    That can be checked using reverse DFS over pairs of states. Finding the
    shortest synchronizing word is NP-complete.
  \subsection{Max flow with lower bounds on edges}
    % TODO: Test this!
    Change edge $(u,v,l\leq f\leq c)$ to $(u,v,f\leq c-l)$. Add edge
    $(t,s,\infty)$. Create super-nodes $S$, $T$. Let $M(u) = \sum_{v}
    l(v,u) - \sum_{v} l(u,v)$. If $M(u)<0$, add edge $(u,T,-M(u))$, else
    add edge $(S,u,M(u))$. Max flow from $S$ to $T$. If all edges from $S$
    are saturated, then we have a feasible flow. Continue running max flow
    from $s$ to $t$ in original graph.
    % TODO: Was there something similar for vertex capacities that we should add?
  \subsection{Tutte matrix for general matching}
    Create an $n\times n$ matrix $A$. For each edge $(i,j)$, $i<j$, let
    $A_{ij} = x_{ij}$ and $A_{ji} = -x_{ij}$. All other entries are $0$.
    The determinant of $A$ is zero iff.\ the graph has a perfect matching.
    A randomized algorithm uses the Schwartz--Zippel lemma to check if it is
    zero.
\subsection{Counting Spanning Trees}
    Kirchoff's Theorem: The number of spanning trees of any graph is the
    determinant of any cofactor of the Laplacian matrix in $O(n^3)$.
    \begin{enumerate}
        \item Let $A$ be the adjacency matrix.
        \item Let $D$ be the degree matrix (matrix with vertex degrees on the diagonal).
        \item Get $D-A$ and delete exactly one row and column. Any row and
        column will do. This will be the cofactor matrix.
        \item Get the determinant of this cofactor matrix using Gauss-Jordan.
        \item $\text{Spanning Trees} = \left|\mathrm{cofactor}(D-A) \right|$
    \end{enumerate}
    
  \subsection{Erd\H{o}s-Gallai Theorem}
    A sequence of non-negative integers $d_1 \ge \cdots \ge d_n$ can be represented as the
    degree sequence of finite simple graph on $n$ vertices if and only if $d_1 + \cdots + d_n$ is
    even and the following holds for $1 \le k \le n$:
    \[
    \sum_{i=1}^n d_i \le k(k-1) + \sum_{i=k+1}^n \min\left(d_i, k\right)
    \]
    
    \subsection{Spherical Coordinates}
    \[
        \begin{array}{cc}
            x = r \cos \theta \cos \phi & r = \sqrt{x^2 + y^2 + z^2} \\
            y = r \cos \theta \sin \phi & \theta = \cos^{-1} x/r \\
            z = r \sin \theta & \phi = \mathrm{atan2}(y,x)
        \end{array}
    \]
    
      \subsectionRed{Delaunay Triangulation}
    Simply map each point $(x,y)$ to $(x,y,x^2+y^2)$, find the 3d convex hull, and drop the 3rd dimension.
    \subsection{Line upper/lower envelope}
    To find the upper/lower envelope of a collection of lines $a_i+b_i x$,
    plot the points $(b_i,a_i)$, add the point $(0,\pm \infty)$ (depending
    on if upper/lower envelope is desired), and then find the convex hull.
  \subsection{Formulas}
    Let $a = (a_x, a_y)$ and $b = (b_x, b_y)$ be two-dimensional vectors.
    \begin{itemize}
      \item $a\cdot b = |a||b|\cos{\theta}$, where $\theta$ is the angle
        between $a$ and $b$.
      \item $a\times b = |a||b|\sin{\theta}$, where $\theta$ is the
        signed angle between $a$ and $b$.
      \item $a\times b$ is equal to the area of the parallelogram with
        two of its sides formed by $a$ and $b$. Half of that is the
        area of the triangle formed by $a$ and $b$.
      \item The line going through $a$ and $b$ is $Ax+By=C$ where $A=b_y-a_y$, $B=a_x-b_x$, $C=Aa_x+Ba_y$.
      \item Two lines $A_1x+B_1y=C_1$, $A_2x+B_2y=C_2$ are parallel iff.\ $D=A_1B_2-A_2B_1$ is zero. Otherwise their unique intersection is $(B_2C_1-B_1C_2,A_1C_2-A_2C_1)/D$.
      \item \textbf{Euler's formula:} $V - E + F = 2$
      \item Side lengths $a,b,c$ can form a triangle iff.\ $a+b>c$, $b+c>a$ and $a+c>b$.
      \item Sum of internal angles of a regular convex $n$-gon is $(n-2)\pi$.
      \item \textbf{Law of sines:} $\frac{a}{\sin A} = \frac{b}{\sin B} = \frac{c}{\sin C}$
      \item \textbf{Law of cosines:} $b^2 = a^2 + c^2 - 2ac\cos B$
      \item Internal tangents of circles $(c_1,r_1), (c_2,r_2)$ intersect at $(c_1r_2+c_2r_1)/(r_1+r_2)$, external intersect at $(c_1r_2-c_2r_1)/(r_1+r_2)$.
    \end{itemize}
    \subsection{Fibonacci Matrix}
    Fast computation for $n$th Fibonacci $\left\{F_1,F_2,\ldots,F_n \right\}$ in $O(\log n)$:
    \[
    \begin{bmatrix}
        F_n \\
        F_{n-1}
    \end{bmatrix}
    =
    \begin{bmatrix}
        1 & 1 \\
        1 & 0
    \end{bmatrix}^n
    \times
    \begin{bmatrix}
        F_2 \\
        F_1
    \end{bmatrix}
    \]
    
    \subsection{Granville's Theorem}
    Compute $\binom{n}{k} \bmod{m}$ (for any $m$) in $O(m^2 \log^2 n)$ time.
    \subsection{Derangements}
    Compute the number of permutations with $n$ elements such that no element is at their original position:
    \[
    D(n) = (n-1) \left( D(n-1) + D(n-2) \right) =  n D(n-1) + (-1)^n
    \]
    % \subsection{Next Permutation}
    % \subsection{Next Combination}
    % \subsection{Next Arrangement}
    % \subsection{Next Partition}
    
      \subsection{Stirling Numbers}
    $s_1$: Count the number of permutations of $n$ elements with $k$ disjoint cycles

    $s_2$: Count the ways to partition a set of $n$ elements into $k$ nonempty subsets
    \[
    s_1(n,k) = \begin{cases}
        1 & n = k = 0 \\
        s_1(n-1,k-1) - (n-1) s_1(n-1,k) & n,k>0 \\
        0 & \text{elsewhere}
    \end{cases}
    \]
    \[
    s_2(n,k) = \begin{cases}
        1 & n = k = 0 \\
        s_2(n-1,k-1) + k s_2(n-1,k) & n,k>0 \\
        0 & \text{elsewhere}
    \end{cases}
    \]
  \subsection{Partition Function}
    Pregenerate the number of partitions of positive integer $n$ with $n$ positive addends.
    \[
    p(n,k) = \begin{cases}
        1 & n = k = 0 \\
        0 & n < k \\
        p(n-1,k-1) + p(n-k,k) & n \ge k
    \end{cases}
    \]
    
  \subsection{128-bit Integer}
    GCC has a 128-bit integer data type named \texttt{\_\_int128}. Useful
    if doing multiplication of 64-bit integers, or something needing a
    little more than 64-bits to represent. There's also
    \texttt{\_\_float128}.
    
      \subsection{Number/Sum of Divisors}
    If a number $n$ is prime factorized where $n = {p_1}^{e_1} \times {p_2}^{e_2} \times \cdots \times {p_k}^{e_k}$, where $\sigma_0$ is the number of divisors while $\sigma_1$ is the sum of divisors:
    \[
    \sum_{d\mid n} d^k = \sigma_k (n) = \prod \frac{{p_i}^{k(e_i)+1}-1}{p_i -1}
    \]
    \[
    \text{Product: } \prod_{d\mid n} d = n^{\frac{\sigma_1 (n)}{2}}
    \]
    
      \subsection{M\"{o}bius Inversion}
    Given arithmetic functions $f$ and $g$:
    \[
    g(n) = \sum_{d\mid n} f(d) \quad \Leftrightarrow \quad f(n) = \sum_{d\mid n} \mu(d)\; g\left(\frac{n}{d}\right)
    \]
    
      \subsection{Number of Integer Points under a Lines}
    Count the number of integer solutions to $Ax+By \leq C$, $0 \leq x \leq n$,
    $0 \leq y$. In other words, evaluate the sum $\sum_{x=0}^n\left\lfloor\dfrac{C-Ax}{B}+1\right\rfloor$.
    To count all solutions, let $n = \left\lfloor\dfrac{c}{a}\right\rfloor$.
    In any case, it must hold that $C-nA \geq 0$. Be very careful about overflows.

